---
title: "Wildfires Competition"
subtitle: "GO Bayes!"
author: "Eric"
date: "27/4/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = NA)
```

# Data loading

First load the data, add a column with site name, and one with a date 
```{r librairies, cache = FALSE}
rm(list=ls())
## ----librairies---------------------------------------------------------------------------------------------------

require(tidyverse) # Manipulation de donnees
require(readxl) # Lecture de fichiers excel
require(lubridate) # Manipulation de dates

## ----gros fichier---------------------------------------------------------------------------------------------------------

load("d1.RData")
load("d2.RData")
load("d3.RData")

data_train_DF=rbind(d1,d2,d3)
rm(d1,d2,d3)
# ajout code site et date
data_train_DF<-data_train_DF %>% 
  unite(col = site, c("lon","lat"),remove = FALSE) %>% 
  mutate(site=str_replace(site,pattern = "-",replacement = "Lon")) %>% 
  mutate(site=str_replace(site,pattern = "_",replacement = "lat")) %>% 
  mutate(date=lubridate::make_datetime(year=year, month=month)) %>% 
  rownames_to_column(var="IdNumber") %>% mutate(IdNumber=as.integer(IdNumber))
```

There are $3503$ sites that are measured during $7$ months during $23$ years, which makes $563983$ records.


And group by site to explore the data
```{r}
d <- data_train_DF %>% group_by(site) %>% 
  summarize(lat=mean(lat),lon=mean(lon),CNTmax=max(CNT, na.rm=T), CNTmean=mean(CNT, na.rm=T),
            BAmax=max(BA, na.rm=T), BAmean=mean(BA, na.rm=T)) %>%
  arrange(desc(CNTmean) ) %>%relocate(lat,lon, .after= BAmean) %>% tibble()
head(d)

d %>% ggplot(aes(y=log(1+CNTmean), x=log(1+BAmean)))+geom_point()+geom_smooth(se=0)
#data_train_DF%>% filter(month %in% c(5,6) ) %>% ggplot(aes(y=log(1+CNT), x=log(1+BA)))+geom_point()+geom_smooth(se=0)
```


# Data formatting

## Binning discrete variables

We recode the number of fires and their damages using their categorical values given by u_ba and u_cnt. As max(u_ba) is geater than max(data_train_DF) we add one extremal category. We now consider 29 bins.
```{r}
#u_baprim=c(-1,u_ba)
#cut(0,u_baprim, labels=F)
#cut(60000,u_baprim, labels=F)
#cut(NA,u_baprim, labels=F)
cut_modif <- function(feux=0,u=u_ba){
  uprim=c(-1,u)
  ifelse(is.na(feux),yes=feux, no= cut(feux,uprim, labels=F))
}

data_train_DF<- data_train_DF %>% 
  mutate(tetha_BA=cut_modif(feux=BA,u=c(u_ba,1000000)),
                    theta_CNT=cut_modif(feux=CNT,u=c(u_cnt,1000))     ) 
data_train_DF %>% pull(theta_CNT) %>% table()->tab
prior_CNT= (tab/sum(tab))
cumsum(prior_CNT)
```

## Spatial coherences

```{r}
decoupe= c(0.0, 0.1, 0.2 ,0.3, 0.4, 0.5, 0.75, 0.9, 0.95, 0.99, 1.0)
data_train_DF %>% group_by(site) %>% summarize(lon=mean(lon), lat=mean(lat) , CNT=mean(CNT, na.rm=T)) %>% pull(CNT) %>% quantile(prob=decoupe) %>% as.numeric->ubis
d<- data_train_DF %>% group_by(site) %>% summarize(lon=mean(lon), lat=mean(lat) , CNT=mean(CNT, na.rm=T), BA=mean(BA, na.rm=T)) %>% mutate(classe_CNT= cut_modif(CNT,u=ubis) ) 
d %>%  ggplot(aes(x=lon, y=lat))+
  geom_tile(aes(fill=-classe_CNT), color="black")+
  theme(legend.position="none") +
  labs(title="Grouping means of Wildfires Counts", x="Longitude", y="Latitude")+
  scale_fill_gradientn(colours = heat.colors(length(decoupe)))
  
```


```{r}
decoupe= c(0.0, 0.1, 0.2 ,0.3, 0.4, 0.5, 0.75, 0.9, 0.95, 0.99, 1.0)
d %>% pull(BA) %>% quantile(prob=decoupe) %>% as.numeric->ubis

d <- d %>% mutate(classe_BA= cut_modif(BA,u=ubis) ) 

d %>%  ggplot(aes(x=lon, y=lat))+
  geom_tile(aes(fill=-classe_BA), color="black")+
  theme(legend.position="none") +
  labs(title="Grouping means of Wildfires Burnt Areas", x="Longitude", y="Latitude")+
  scale_fill_gradientn(colours = heat.colors(length(decoupe)))
  
```

Adding the classes for BA and CNT to the main file

```{r}
d<-d %>% rename(mean_CNT=CNT, mean_BA=BA,
                classe_meanCNT=classe_CNT,
                classe_meanBA=classe_BA)
data_train_DF<-data_train_DF %>% left_join(d)
rm(d)
```

## Looking for spatial neighbors

We create variables nbcolat and nbcolong that say how may sites can be found with the same latitude (or with the same longitude) at a given time.
```{r}
d <- data_train_DF %>% dplyr::select(IdNumber,site,lat,lon,date,theta_CNT,tetha_BA)
d<-d %>% arrange(date,lon,lat)
d %>% group_by(date,lon) %>% summarise(nbcolong=length(lat)) %>% 
  ungroup() -> dcolong
d %>% group_by(date,lat) %>% summarise(nbcolat=length(lon)) %>% 
  ungroup() -> dcolat
d<-d %>% left_join(dcolat) %>% left_join(dcolong)
```

We first arrange by date and longitude followed by latitude. There is 7 months $\times$ 23 years $\times$ 117 = 18837 blocs of different latitudes. In this setting, one goes from left to right (following increasing longitudes) and then from bottom to top (following increasing latitudes). The regular structure of the lattice allows for a quick evaluation of the closest neighbors belonging to the same longitude.
```{r}
d<-d %>% arrange(date,lon,lat)%>% rowid_to_column(var="IdLat") 
nsites <- 3503
delta_up<- delta_bottom <- rep(0,nsites)
e=1
repeat {
colong=d$nbcolong[e]
delta_bottom[e]=0
delta_up[e+colong-1]= 0
if(colong!=1){
delta_bottom[(e+1):(e+colong-1)]=-1
delta_up[(e):(e+colong-2)]= 1
}
e=e+colong
if (e >=(nsites+1)){
break
}
}
d$IdNumber_down=d$IdNumber[d$IdLat+rep(delta_bottom, 161)]  
d$IdNumber_up=d$IdNumber[d$IdLat+rep(delta_up, 161)] 

# a bad computational idea is as follows
# E <- length(d$lat)
# right_neigbor<-rep(NA,E)
# left_neigbor<-rep(NA,E)
# up_neigbor<-rep(NA,E)
# down_neigbor<-rep(NA,E)
# for(e in 1:E) {
#   lat<-d$lat[e]
#   long<-d$lon[e]
#   dat=d$date[e]
#   e_right<-which((d$date==dat)&(d$lat==lat)&(d$long==long+0.5))
#   e_left<-which((d$date==dat)&(d$lat==lat)&(d$long==long-0.5))
#   e_up<-which((d$date==dat)&(d$lat==lat+0.5)&(d$long==long))
#   e_down<-which((d$date==dat)&(d$lat==lat-0.5)&(d$long==long))
#   
#   right_neigbor[e]=ifelse(length(e_right)!=0,e_right,NA )
#   left_neigbor[e]= ifelse(length(e_left)!=0,e_left,NA )
#    up_neigbor[e]= ifelse(length(e_up)!=0,e_up,NA )
#     down_neigbor[e]=ifelse(length(e_down)!=0,e_down,NA )
#     
# 
# }
```

We can work the same way with neighbors belonging to the same latitude
```{r}
d<-d %>% arrange(date,lat,lon)%>% rowid_to_column(var="IdLong") 
nsites <- 3503
delta_left<- delta_right <- rep(0,nsites)
e=1
repeat {
colat=d$nbcolat[e]
delta_left[e]=0
delta_right[e+colong-1]= 0
if(colat!=1){
delta_left[(e+1):(e+colat-1)]=-1
delta_right[(e):(e+colat-2)]= 1
}
e=e+colat
if (e >=(nsites+1)){
break
}
}
d$IdNumber_left=d$IdNumber[d$IdLong+rep(delta_left, 161)]  
d$IdNumber_right=d$IdNumber[d$IdLong+rep(delta_right, 161)] 


data_train_DF<- data_train_DF %>% left_join(d %>% dplyr::select(starts_with("IdNumber")))
```

One can check that the neighbors seem correctly identified...

```{r}
sixselections <-sample(data_train_DF$IdNumber,6,rep=F)
ens=NULL
for(s in sixselections){ens=c(ens,data_train_DF %>% filter(IdNumber==s) %>% dplyr::select(starts_with("IdNum")) %>% as.vector() ) }
d<-data_frame(IdNumber=as.numeric(ens),coul=rep(1:6,each=5)) %>% left_join(data_train_DF %>% dplyr::select(IdNumber,lat,lon))
data_train_DF %>%  ggplot(aes(x=lon, y=lat))+
  geom_tile(color="black", fill="white")+geom_tile(data=d,fill=as.factor(d$coul))
```


```{r}
check_neighbor <- function(id,voisin,V_id,V_voisin){
  ifelse(abs(V_id-V_voisin) > 0.51, yes=id, no=voisin)
}
# voisin<-data_train_DF$IdNumber_left
# V_voisin= data_train_DF$lon[voisin]
# id=data_train_DF$IdNumber
# V_id=data_train_DF$lon
#sum(check_neighbor(id,voisin,V_id,V_voisin) != voisin)

data_train_DF<- data_train_DF %>%
  mutate(IdNumber_left=check_neighbor(IdNumber,IdNumber_left,lon,lon[IdNumber_left])) %>% 
  mutate(IdNumber_right=check_neighbor(IdNumber,IdNumber_right,lon,lon[IdNumber_right])) %>%        
  mutate(IdNumber_down=check_neighbor(IdNumber,IdNumber_down,lat,lat[IdNumber_down])) %>%
  mutate(IdNumber_left=check_neighbor(IdNumber,IdNumber_up,lat,lat[IdNumber_up])) 
  
```

## Looking for temporal neighbors

```{r}
data_train_DF %>% group_by(date,classe_meanCNT) %>% summarise(CNT=mean(CNT,na.rm=T)) %>%  ggplot(aes(x=date,y=log(CNT+1), color=as.factor(classe_meanCNT) ))+
  geom_line()+geom_smooth(se=0)+ theme(legend.position="none") +
  labs(title="Time evolution of Wildfires Counts", x="Date")

data_train_DF %>% group_by(month,classe_meanBA) %>% summarise(CNT=mean(CNT,na.rm=T)) %>%  ggplot(aes(x=month,y=CNT, color=as.factor(classe_meanBA) ))+
  geom_line()+ theme(legend.position="none") +
  labs(title="Monthly evolution of mean Wildfires Counts by classes of Burnt Areas", x="Month")
```


```{r}
data_train_DF %>% group_by(date,classe_meanBA) %>% summarise(BA=mean(BA,na.rm=T)) %>%  ggplot(aes(x=date,y=log(BA+1), color=as.factor(classe_meanBA) ))+
  geom_line()+geom_smooth(se=0)+ theme(legend.position="none") +
  labs(title="Time evolution of Wildfires Burnt areas", x="Date")

data_train_DF %>% group_by(month,classe_meanCNT) %>% summarise(BA=mean(BA,na.rm=T)) %>%  ggplot(aes(x=month,y=BA, color=as.factor(classe_meanCNT) ))+
  geom_line()+ theme(legend.position="none") +
  labs(title="Monthly evolution of mean Wildfires Burnt Areas by classes of Fire Counts", x="Month")
```

We now try to get the previous time record at a given spatio-temporal record, as well as the next one. If it appears to be the first month in the year the spatio-temporal record is its own previous neighbour. The point is its own next time neighbour if it correspond to the last month in the year.

```{r}
d <- data_train_DF %>% dplyr::select(IdNumber,site,lat,lon,date,month,year)
d<-d %>% arrange(lon,lat,date) %>% rowid_to_column(var="IdTime") %>% 
  mutate(IdTime=as.integer(IdTime))

delta_next <-rep(c(rep(1,each=6),0),each=80569)
d$IdNumber_next=d$IdNumber[d$IdTime+delta_next]  

delta_previous <- rep(c(0,rep(-1,each=6)),each=80569)
d$IdNumber_previous=d$IdNumber[d$IdTime+delta_previous]  

data_train_DF<- data_train_DF %>% left_join(d %>% dplyr::select(starts_with("IdNumber")))



```

# Test and Training Data

```{r}
 # Completer data_train_DF par qualitatives
data_train_DF<- data_train_DF %>%mutate(theta_CNT_left=theta_CNT[IdNumber_left],
                                        theta_CNT_right=theta_CNT[IdNumber_right],
                                        theta_CNT_up=theta_CNT[IdNumber_up],
                                        theta_CNT_down=theta_CNT[IdNumber_down],
                                        theta_CNT_previous=theta_CNT[IdNumber_previous],
                                        theta_CNT_next=theta_CNT[IdNumber_next])

 # Completer data_train_DF par quantitativesNQT on LC
LC<- data_train_DF %>% dplyr::select(.,starts_with("lc")) %>% as.matrix()
for(j in 1:18){
  Fn<-ecdf(LC[,j])
LC[,j]= qnorm(0.99*Fn(LC[,j]))
}

ACP<-princomp(LC,cor=TRUE,scores=TRUE)
LC=ACP$scores
colnames(LC)=paste("lc",1:18,sep="")

data_train_DF<- cbind(data_train_DF %>% dplyr::select(.,-starts_with("lc")),LC)
 
# Completer data_train_DF par quantitativesNQT on LC
CLIM<- data_train_DF %>% dplyr::select(.,starts_with("clim")) %>% as.matrix()
for(j in 1:10){
  Fn<-ecdf(CLIM[,j])
CLIM[,j]= qnorm(0.99*Fn(CLIM[,j]))
}

ACP<-princomp(CLIM,cor=TRUE,scores=TRUE)
CLIM=ACP$scores
colnames(CLIM)=paste("clim",1:10,sep="")
data_train_DF<- cbind(data_train_DF %>% dplyr::select(.,-starts_with("clim")),CLIM)
 


 # Selection apprentissage et test
train_CNT_Id <-data_train_DF %>% filter(!is.na(CNT)) %>% pull(IdNumber)
#sum(train_CNT_Id[-1] < train_CNT_Id[-length(train_CNT_Id)])
test_CNT_Id <-data_train_DF %>% filter(is.na(CNT)) %>% pull(IdNumber)
#sum(test_CNT_Id[-1] < test_CNT_Id[-length(test_CNT_Id)])
length(train_CNT_Id)
length(test_CNT_Id)

test<-data_train_DF %>% filter(is.na(CNT))
tada<-data_train_DF %>% filter(!is.na(CNT))

test_select_Id<-test %>% pull(IdNumber)
train_select_Id<-tada  %>% pull(IdNumber)

length(test_select_Id)
length(train_select_Id)

```
The IdNumber column is sorted in dataframes test and tada


# Bayesian Updating

## Bayesian updating in theory


Let's call $\theta$ the unknown, $[\theta]$ its probability distribution.
Consider some data $(A,B)$ taken as pieces of information. Via Bayes'rule one gets
$$[\theta|A,B]=\frac{[\theta,A,B]}{[A,B]}=\frac{[A,B|\theta]\times [\theta]}{[A,B]}\\
[\theta|A,B]=\frac{[A|\theta,B]\times[B|\theta]\times [\theta]}{[A,B]}$$
Focusing on these quantities as a function of the target $\theta$ only, we see that:
$$[{\theta}|A,B] \propto [A|\theta,B]\times[B|\theta]\times [\theta]$$
In the challenge, $\theta$ is a categorical variable refering eiter to the count class of wildfires (ranging from $1$ to $28$) in a given pixel at a certain time, or the corresponding burnt area or both. The available information $A$ stems from many various sources:
 - the neigbours because of the spatial coherence,
 - the previous category of the pixel (or its neighbours), as well as its next state due to some expected temporal behavior.
 - the 10 meteorological variables.
 - the proportions of the 18 land cover classes in the grid cell.
 When the information $A$ and the unknown $\theta$ are both categorical as in the first two cases, $A|\theta$ is a cross-table that can be empirically computed from the raw data table.
 When the information $A$ is continuous and the unknown $\theta$ fixed at a certain category, $A|\theta$ empirically correspond to a matrix of observations that could be modeled as a multivariate Gaussian realisation after suitable Normal Quantile Transformations.
 
 Assuming $[A,B|\theta]=[A|\theta]\times [B|\theta]$ i.e. that, given $\theta$, the pieces of information $A$ and $B$ are conditionnally independent would allow for sequential updating:
$$[{\theta}|A,B] \propto
[A|\theta]\times[B|\theta]\times [\theta]$$

The following function computes the score for the competition. Mind that only the 28 components are needed!


```{r}
SCNT_cat=function(Obs, Pred,u_cnt=u_cnt,weights_cnt=weights_cnt) {
  
  # calculate the matrix with estimated exceedance probability of the severity thresholds:
  indicatrice_cnt=Pred
  prediction_cnt=t(apply(Pred,1, cumsum))
  
  for(k in 1:length(u_cnt)){
    Ind_k=Obs
    Ind_k[u_cnt[k]<Obs]<-0
    Ind_k[u_cnt[k]>=Obs]<-1
    indicatrice_cnt[,k]=Ind_k
  }
  
  #Computation of Scnt
#  weights_cnt=matrix(weights_cnt,ncol=1)
#  Scnt=sum(((indicatrice_cnt-prediction_cnt)^2%*%weights_cnt))
  Scnt=sum((as.matrix((indicatrice_cnt-prediction_cnt)^2)%*%matrix(weights_cnt,ncol=1)))
  
  return(Scnt)}


```


 We start with a prior based on observed frequencies
 
```{r}
prior_CNT = as.numeric(table(tada$theta_CNT)/sum(table(tada$theta_CNT)))
```
 
 We first update with information stemming from the class of mean number of fires 
 
```{r}
InfoSachantTheta=function(d=tibble(IdNumber=rep(0,11*29),
                                       Info=rep(1:11,29),
                                       Theta=rep(1:29 ,each=11)),
                                   n_Info = 11,
                                   n_Theta=29){
  truc= tibble(IdNumber=rep(0,n_Info*n_Theta),
               Info=rep(1:n_Info,n_Theta),
               Theta=rep(1:n_Theta ,each=n_Info))
  dd<- rbind(d,truc)
  TAB= as.matrix(table(dd$Theta,dd$Info))-1
  res=TAB/apply(TAB,1,sum)
}



MatInfoSachantTheta= InfoSachantTheta(d=tada %>% dplyr::select(IdNumber,theta_CNT,classe_meanCNT) %>% 
                                        rename(Theta=theta_CNT, Info=classe_meanCNT),
                                      n_Info = 11,
                                      n_Theta=29)
  
BayesUpdate=function(MatInfoSachantTheta, VectInfo, MatProba){
  AsachantT=0*MatProba
  AsachantT=MatInfoSachantTheta[,VectInfo]
  TsachantA=t(AsachantT)*MatProba
  TsachantA=TsachantA/apply(TsachantA,1,sum)
}

Proba= BayesUpdate(MatInfoSachantTheta= MatInfoSachantTheta,
            VectInfo = test$classe_meanCNT, MatProba= matrix(as.numeric(prior_CNT),byrow=T,nrow=length(test$classe_meanCNT),ncol=29))

score= list(init=
              SCNT_cat(Obs=test$CNT, Pred=(matrix(as.numeric(prior_CNT[1:28]),byrow=T,nrow=length(test$CNT),ncol=28)) ,u_cnt=u_cnt,weights_cnt=weights_cnt)
           )
score= c(score, list( avecMoyCNT= SCNT_cat(Obs=test$CNT, Pred= Proba[,1:28] ,u_cnt=u_cnt,weights_cnt=weights_cnt)
        ))
(score)

colnames(Proba)=paste("p",1:29,sep="")
Proba=as.data.frame.matrix(Proba)
head(Proba)

test=cbind(test,Proba)

```
 
 
## Bayesian updating in practice


We now update the probability of the CNT states with neighbors located at the left

```{r}
Proba0= test %>% dplyr::select(starts_with("p"))
MatInfoSachantTheta= InfoSachantTheta(d=tada %>% 
                                        dplyr::select(IdNumber,IdNumber_left,
                                                      theta_CNT,theta_CNT_left) %>% 
                                        filter(IdNumber_left %in% train_CNT_Id) %>% 
                                        filter(IdNumber_left!=IdNumber) %>% 
                                        rename(Theta=theta_CNT, Info= theta_CNT_left) %>% 
                                        dplyr::select(IdNumber,Theta,Info) ,
                                      n_Info = 29,
                                      n_Theta=29)

OK <- test %>% filter(IdNumber_left!=IdNumber) %>% 
               filter(!is.na(theta_CNT_left))

Nope <- test %>% filter(!(IdNumber %in% OK$IdNumber))

ProbaOK= BayesUpdate(MatInfoSachantTheta= MatInfoSachantTheta,
            VectInfo = OK$theta_CNT_left, 
            MatProba = OK %>% 
              dplyr::select(starts_with("p")))

ProbaOK<-ProbaOK %>% mutate(IdNumber=OK$IdNumber)
ProbaNope= Nope %>% dplyr::select(starts_with("p"),IdNumber)
test=test %>% dplyr::select(-starts_with("p")) %>% 
  left_join(rbind(ProbaOK,ProbaNope)) %>% arrange(IdNumber)
Proba= test %>% dplyr::select(starts_with("p"))
perf_left<- SCNT_cat(Obs=test$CNT, Pred= Proba[,1:28] ,u_cnt=u_cnt,weights_cnt=weights_cnt)
score<-c(score, list(left=perf_left))

#l<-ls()
#rm(list=l[!l%in% c("data_train_DF", "Proba","u_ba","u_cnt","weights_ba","weights_cnt","SCNT_cat","scores","CalculProbaDiscrete")])
```

We now update the probability of the CNT states with neighbors located at the right

```{r}
Proba0= test %>% dplyr::select(starts_with("p"))
MatInfoSachantTheta= InfoSachantTheta(d=tada %>% 
                                        dplyr::select(IdNumber,IdNumber_right,
                                                      theta_CNT,theta_CNT_right) %>% 
                                        filter(IdNumber_right %in% train_CNT_Id) %>% 
                                        filter(IdNumber_right!=IdNumber) %>% 
                                        rename(Theta=theta_CNT, Info= theta_CNT_right) %>% 
                                        dplyr::select(IdNumber,Theta,Info) ,
                                      n_Info = 29,
                                      n_Theta=29)
OK <- test %>% filter(IdNumber_right!=IdNumber) %>% 
               filter(!is.na(theta_CNT_right))

Nope <- test %>% filter(!(IdNumber %in% OK$IdNumber))

ProbaOK= BayesUpdate(MatInfoSachantTheta= MatInfoSachantTheta,
            VectInfo = OK$theta_CNT_right, 
            MatProba = OK %>% 
              dplyr::select(starts_with("p")))
ProbaOK<-ProbaOK %>% mutate(IdNumber=OK$IdNumber)
ProbaNope= Nope %>% dplyr::select(starts_with("p"),IdNumber)
test=test %>% dplyr::select(-starts_with("p")) %>% 
  left_join(rbind(ProbaOK,ProbaNope)) %>% arrange(IdNumber)
Proba= test %>% dplyr::select(starts_with("p"))
perf_right<- SCNT_cat(Obs=test$CNT, Pred= Proba[,1:28] ,u_cnt=u_cnt,weights_cnt=weights_cnt)
score<-c(score, list(right=perf_right))

##rm(list=l[!l%in% c("data_train_DF", "Proba","u_ba","u_cnt","weights_ba","weights_cnt","SCNT_cat","scores","CalculProbaDiscrete")])
```

With neighbors located North

```{r}
Proba0= test %>% dplyr::select(starts_with("p"))
MatInfoSachantTheta= InfoSachantTheta(d=tada %>%
                                        dplyr::select(IdNumber,IdNumber_up,
                                                      theta_CNT,theta_CNT_up) %>%
                                        filter(IdNumber_up %in% train_CNT_Id) %>%
                                        filter(IdNumber_up!=IdNumber) %>%
                                        rename(Theta=theta_CNT, Info= theta_CNT_up) %>%
                                        dplyr::select(IdNumber,Theta,Info) ,
                                      n_Info = 29,
                                      n_Theta=29)
OK <- test %>% filter(IdNumber_up!=IdNumber) %>%
               filter(!is.na(theta_CNT_up))

Nope <- test %>% filter(!(IdNumber %in% OK$IdNumber))

ProbaOK= BayesUpdate(MatInfoSachantTheta= MatInfoSachantTheta,
            VectInfo = OK$theta_CNT_up,
            MatProba = OK %>%
              dplyr::select(starts_with("p")))
ProbaOK<-ProbaOK %>% mutate(IdNumber=OK$IdNumber)
ProbaNope= Nope %>% dplyr::select(starts_with("p"),IdNumber)
test=test %>% dplyr::select(-starts_with("p")) %>%
left_join(rbind(ProbaOK,ProbaNope)) %>% arrange(IdNumber)
Proba= test %>% dplyr::select(starts_with("p"))
perf_up<- SCNT_cat(Obs=test$CNT, Pred= Proba[,1:28] ,u_cnt=u_cnt,weights_cnt=weights_cnt)
score<-c(score, list(up_pas_pris=perf_up))
test = cbind(test %>% dplyr::select(-starts_with("p")),Proba0)



```

With neighbors located South

```{r}
Proba0= test %>% dplyr::select(starts_with("p"))
MatInfoSachantTheta= InfoSachantTheta(d=tada %>% 
                                        dplyr::select(IdNumber,IdNumber_down,
                                                      theta_CNT,theta_CNT_down) %>% 
                                        filter(IdNumber_down %in% train_CNT_Id) %>% 
                                        filter(IdNumber_down!=IdNumber) %>% 
                                        rename(Theta=theta_CNT, Info= theta_CNT_down) %>% 
                                        dplyr::select(IdNumber,Theta,Info) ,
                                      n_Info = 29,
                                      n_Theta=29)
OK <- test %>% filter(IdNumber_down!=IdNumber) %>% 
               filter(!is.na(theta_CNT_down))

Nope <- test %>% filter(!(IdNumber %in% OK$IdNumber))

ProbaOK= BayesUpdate(MatInfoSachantTheta= MatInfoSachantTheta,
            VectInfo = OK$theta_CNT_down, 
            MatProba = OK %>% 
              dplyr::select(starts_with("p")))
ProbaOK<-ProbaOK %>% mutate(IdNumber=OK$IdNumber)
ProbaNope= Nope %>% dplyr::select(starts_with("p"),IdNumber)
test=test %>% dplyr::select(-starts_with("p")) %>% left_join(rbind(ProbaOK,ProbaNope))
Proba= test %>% dplyr::select(starts_with("p"))
perf_down<- SCNT_cat(Obs=test$CNT, Pred= Proba[,1:28] ,u_cnt=u_cnt,weights_cnt=weights_cnt)
score<-c(score, list(down=perf_down))

```

With neighbours occuring before

```{r}
Proba0= test %>% dplyr::select(starts_with("p"))
MatInfoSachantTheta= InfoSachantTheta(d=tada %>% 
                                        dplyr::select(IdNumber,IdNumber_previous,
                                                      theta_CNT,theta_CNT_previous) %>% 
                                        filter(IdNumber_previous %in% train_CNT_Id) %>% 
                                        filter(IdNumber_previous!=IdNumber) %>% 
                                        rename(Theta=theta_CNT, Info= theta_CNT_previous) %>% 
                                        dplyr::select(IdNumber,Theta,Info) ,
                                      n_Info = 29,
                                      n_Theta=29)
OK <- test %>% filter(IdNumber_previous!=IdNumber) %>% 
               filter(!is.na(theta_CNT_previous))

Nope <- test %>% filter(!(IdNumber %in% OK$IdNumber))

ProbaOK= BayesUpdate(MatInfoSachantTheta= MatInfoSachantTheta,
            VectInfo = OK$theta_CNT_previous, 
            MatProba = OK %>% 
              dplyr::select(starts_with("p")))
ProbaOK<-ProbaOK %>% mutate(IdNumber=OK$IdNumber)
ProbaNope= Nope %>% dplyr::select(starts_with("p"),IdNumber)
test=test %>% dplyr::select(-starts_with("p")) %>% left_join(rbind(ProbaOK,ProbaNope))
Proba= test %>% dplyr::select(starts_with("p"))
perf_previous<- SCNT_cat(Obs=test$CNT, Pred= Proba[,1:28] ,u_cnt=u_cnt,weights_cnt=weights_cnt)
score<-c(score, list(previous=perf_previous))
(score)
```
With neighbours occuring after

```{r}
Proba0= test %>% dplyr::select(starts_with("p"))
MatInfoSachantTheta= InfoSachantTheta(d=tada %>% 
                                        dplyr::select(IdNumber,IdNumber_next,
                                                      theta_CNT,theta_CNT_next) %>% 
                                        filter(IdNumber_next %in% train_CNT_Id) %>% 
                                        filter(IdNumber_next!=IdNumber) %>% 
                                        rename(Theta=theta_CNT, Info= theta_CNT_next) %>% 
                                        dplyr::select(IdNumber,Theta,Info) ,
                                      n_Info = 29,
                                      n_Theta=29)
OK <- test %>% filter(IdNumber_next!=IdNumber) %>% 
               filter(!is.na(theta_CNT_next))

Nope <- test %>% filter(!(IdNumber %in% OK$IdNumber))

ProbaOK= BayesUpdate(MatInfoSachantTheta= MatInfoSachantTheta,
            VectInfo = OK$theta_CNT_next, 
            MatProba = OK %>% 
              dplyr::select(starts_with("p")))
ProbaOK<-ProbaOK %>% mutate(IdNumber=OK$IdNumber)
ProbaNope= Nope %>% dplyr::select(starts_with("p"),IdNumber)
test=test %>% dplyr::select(-starts_with("p")) %>% left_join(rbind(ProbaOK,ProbaNope))
Proba= test %>% dplyr::select(starts_with("p"))
perf_next<- SCNT_cat(Obs=test$CNT, Pred= Proba[,1:28] ,u_cnt=u_cnt,weights_cnt=weights_cnt)
score<-c(score, list(after=perf_next))
(score)
```

Save for competition

```{r}
Proba <-test %>% dplyr::select(starts_with("p"))
prediction_cnt = matrix(nrow = 80000, ncol = 28)
prediction_cnt[,1]=Proba[,1]
for(j in 2:28){
  prediction_cnt[,j] = prediction_cnt[,j-1]+Proba[,j]
}
save(list="prediction_cnt", file="NaiveTom.Rdata")
```



