---
title: "Competition"
subtitle: "GO Bayes!"
author: "Eric"
date: "3/3/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = NA)
```

# Data loading

First load the data, add a column with site name, and one with a date 
```{r librairies, cache = FALSE}
rm(list=ls())
source("code_chargement_donnees.R")
```

There are $3503$ sites that are measured during $7$ months during $23$ years, which makes $563983$ records.


And group by site to explore the data
```{r}
data_train_DF %>% group_by(site) %>% 
  summarize(lat=mean(lat),lon=mean(lon),CNTmax=max(CNT, na.rm=T), CNTmean=mean(CNT, na.rm=T),
            BAmax=max(BA, na.rm=T), BAmean=mean(BA, na.rm=T)) %>%
  arrange(desc(CNTmean) ) %>%relocate(lat,lon, .after= BAmean) %>% tibble()

```


# Data formatting

## Binning discrete variables

We recode the number of fires and their damages using their categorical values given by u_ba and u_cnt.
```{r}
#u_baprim=c(-1,u_ba)
#cut(0,u_baprim, labels=F)
#cut(60000,u_baprim, labels=F)
#cut(NA,u_baprim, labels=F)
cut_modif <- function(feux=0,u=u_ba){
  uprim=c(-1,u)
  ifelse(is.na(feux),yes=feux, no= cut(feux,uprim, labels=F))
}

data_train_DF<- data_train_DF %>% mutate(tetha_BA=cut_modif(feux=BA,u=u_ba),
                    theta_CNT=cut_modif(feux=CNT,u=u_cnt)     ) 
data_train_DF %>% pull(theta_CNT) %>% table()->tab
prior_CNT= cumsum(tab/sum(tab))
prior_CNT
```

## Spatial coherences

```{r}
decoupe= c(0.0, 0.1, 0.2 ,0.3, 0.4, 0.5, 0.75, 0.9, 0.95, 0.99, 1.0)
data_train_DF %>% group_by(site) %>% summarize(lon=mean(lon), lat=mean(lat) , CNT=mean(CNT, na.rm=T)) %>% pull(CNT) %>% quantile(prob=decoupe) %>% as.numeric->ubis
d<- data_train_DF %>% group_by(site) %>% summarize(lon=mean(lon), lat=mean(lat) , CNT=mean(CNT, na.rm=T), BA=mean(BA, na.rm=T)) %>% mutate(classe_CNT= cut_modif(CNT,u=ubis) ) 
d %>%  ggplot(aes(x=lon, y=lat))+
  geom_tile(aes(fill=-classe_CNT), color="black")+
  theme(legend.position="none") +
  labs(title="Grouping means of Wildfires Counts", x="Longitude", y="Latitude")+
  scale_fill_gradientn(colours = heat.colors(length(decoupe)))
  
```


```{r}
decoupe= c(0.0, 0.1, 0.2 ,0.3, 0.4, 0.5, 0.75, 0.9, 0.95, 0.99, 1.0)
d %>% pull(BA) %>% quantile(prob=decoupe) %>% as.numeric->ubis

d <- d %>% mutate(classe_BA= cut_modif(BA,u=ubis) ) 

d %>%  ggplot(aes(x=lon, y=lat))+
  geom_tile(aes(fill=-classe_BA), color="black")+
  theme(legend.position="none") +
  labs(title="Grouping means of Wildfires Damages", x="Longitude", y="Latitude")+
  scale_fill_gradientn(colours = heat.colors(length(decoupe)))
  
```

Adding the classes for BA and CNT to the main file

```{r}
d<-d %>% rename(mean_CNT=CNT, mean_BA=BA)
data_train_DF<-data_train_DF %>% left_join(d)
```

## Time evolution

```{r}
data_train_DF %>% group_by(date,classe_CNT) %>% summarise(CNT=mean(CNT,na.rm=T)) %>%  ggplot(aes(x=date,y=log(CNT+1), color=as.factor(classe_CNT) ))+
  geom_line()+geom_smooth(se=0)+ theme(legend.position="none") +
  labs(title="Time evolution of Wildfires Counts", x="Date")
```

## Normal quantile transform

```{r}
order(c(1,NA,3,3,2))


```



