---
title: "Competition"
subtitle: "GO Bayes!"
author: "Eric"
date: "3/3/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = NA)
```

# Data loading

First load the data, add a column with site name, and one with a date 
```{r librairies, cache = FALSE}
rm(list=ls())
source("code_chargement_donnees.R")
```

There are $3503$ sites that are measured during $7$ months during $23$ years, which makes $563983$ records.


And group by site to explore the data
```{r}
data_train_DF %>% group_by(site) %>% 
  summarize(lat=mean(lat),lon=mean(lon),CNTmax=max(CNT, na.rm=T), CNTmean=mean(CNT, na.rm=T),
            BAmax=max(BA, na.rm=T), BAmean=mean(BA, na.rm=T)) %>%
  arrange(desc(CNTmean) ) %>%relocate(lat,lon, .after= BAmean) %>% tibble()

```


# Data formatting

## Binning discrete variables

We recode the number of fires and their damages using their categorical values given by u_ba and u_cnt.
```{r}
#u_baprim=c(-1,u_ba)
#cut(0,u_baprim, labels=F)
#cut(60000,u_baprim, labels=F)
#cut(NA,u_baprim, labels=F)
cut_modif <- function(feux=0,u=u_ba){
  uprim=c(-1,u)
  ifelse(is.na(feux),yes=feux, no= cut(feux,uprim, labels=F))
}

data_train_DF<- data_train_DF %>% mutate(tetha_BA=cut_modif(feux=BA,u=u_ba),
                    theta_CNT=cut_modif(feux=CNT,u=u_cnt)     ) 
data_train_DF %>% pull(theta_CNT) %>% table()->tab
prior_CNT= cumsum(tab/sum(tab))
prior_CNT
```

## Spatial coherences

```{r}
decoupe= c(0.0, 0.1, 0.2 ,0.3, 0.4, 0.5, 0.75, 0.9, 0.95, 0.99, 1.0)
data_train_DF %>% group_by(site) %>% summarize(lon=mean(lon), lat=mean(lat) , CNT=mean(CNT, na.rm=T)) %>% pull(CNT) %>% quantile(prob=decoupe) %>% as.numeric->ubis
d<- data_train_DF %>% group_by(site) %>% summarize(lon=mean(lon), lat=mean(lat) , CNT=mean(CNT, na.rm=T), BA=mean(BA, na.rm=T)) %>% mutate(classe_CNT= cut_modif(CNT,u=ubis) ) 
d %>%  ggplot(aes(x=lon, y=lat))+
  geom_tile(aes(fill=-classe_CNT), color="black")+
  theme(legend.position="none") +
  labs(title="Grouping means of Wildfires Counts", x="Longitude", y="Latitude")+
  scale_fill_gradientn(colours = heat.colors(length(decoupe)))
  
```


```{r}
decoupe= c(0.0, 0.1, 0.2 ,0.3, 0.4, 0.5, 0.75, 0.9, 0.95, 0.99, 1.0)
d %>% pull(BA) %>% quantile(prob=decoupe) %>% as.numeric->ubis

d <- d %>% mutate(classe_BA= cut_modif(BA,u=ubis) ) 

d %>%  ggplot(aes(x=lon, y=lat))+
  geom_tile(aes(fill=-classe_BA), color="black")+
  theme(legend.position="none") +
  labs(title="Grouping means of Wildfires Damages", x="Longitude", y="Latitude")+
  scale_fill_gradientn(colours = heat.colors(length(decoupe)))
  
```

Adding the classes for BA and CNT to the main file

```{r}
d<-d %>% rename(mean_CNT=CNT, mean_BA=BA)
data_train_DF<-data_train_DF %>% left_join(d)
```

## Look for neighbors

We create variables nbcolat and nbcolong that say how may sites can be found with the same latitude (or with the same longitude) at a given time.
```{r}
d <- data_train_DF %>% dplyr::select(IdNumber,site,lat,lon,date,theta_CNT,tetha_BA)
d<-d %>% arrange(date,lon,lat)
d %>% group_by(date,lon) %>% summarise(nbcolong=length(lat)) %>% 
  ungroup() -> dcolong
d %>% group_by(date,lat) %>% summarise(nbcolat=length(lon)) %>% 
  ungroup() -> dcolat
d<-d %>% left_join(dcolat) %>% left_join(dcolong)
```

We first arrange by date and longitude followed by latitude. There is 7 months $\times$23 years$\times$ 117 = 18837 blocs of different latitudes. In this setting, one goes from left to right (following increasing longitudes) and then from bottom to top (following increasing latitudes). The regular structure of the lattice allows for a quick evaluation of the closest neighbors belonging to the same longitude.
```{r}
d<-d %>% arrange(date,lon,lat)%>% rowid_to_column(var="IdLat") 
nsites <- 3503
delta_up<- delta_bottom <- rep(0,nsites)
e=1
repeat {
colong=d$nbcolong[e]
delta_bottom[e]=0
delta_up[e+colong-1]= 0
if(colong!=1){
delta_bottom[(e+1):(e+colong-1)]=-1
delta_up[(e):(e+colong-2)]= 1
}
e=e+colong
if (e >=(nsites+1)){
break
}
}
d$IdNumber_down=d$IdNumber[d$IdLat+rep(delta_bottom, 161)]  
d$IdNumber_up=d$IdNumber[d$IdLat+rep(delta_up, 161)] 

# a bad computational idea is as follows
# E <- length(d$lat)
# right_neigbor<-rep(NA,E)
# left_neigbor<-rep(NA,E)
# up_neigbor<-rep(NA,E)
# down_neigbor<-rep(NA,E)
# for(e in 1:E) {
#   lat<-d$lat[e]
#   long<-d$lon[e]
#   dat=d$date[e]
#   e_right<-which((d$date==dat)&(d$lat==lat)&(d$long==long+0.5))
#   e_left<-which((d$date==dat)&(d$lat==lat)&(d$long==long-0.5))
#   e_up<-which((d$date==dat)&(d$lat==lat+0.5)&(d$long==long))
#   e_down<-which((d$date==dat)&(d$lat==lat-0.5)&(d$long==long))
#   
#   right_neigbor[e]=ifelse(length(e_right)!=0,e_right,NA )
#   left_neigbor[e]= ifelse(length(e_left)!=0,e_left,NA )
#    up_neigbor[e]= ifelse(length(e_up)!=0,e_up,NA )
#     down_neigbor[e]=ifelse(length(e_down)!=0,e_down,NA )
#     
# 
# }
```

We can work the same with neighbors belonging to the same latitude
```{r}
d<-d %>% arrange(date,lat,lon)%>% rowid_to_column(var="IdLong") 
nsites <- 3503
delta_left<- delta_right <- rep(0,nsites)
e=1
repeat {
colat=d$nbcolat[e]
delta_left[e]=0
delta_right[e+colong-1]= 0
if(colat!=1){
delta_left[(e+1):(e+colat-1)]=-1
delta_right[(e):(e+colat-2)]= 1
}
e=e+colat
if (e >=(nsites+1)){
break
}
}
d$IdNumber_left=d$IdNumber[d$IdLong+rep(delta_left, 161)]  
d$IdNumber_right=d$IdNumber[d$IdLong+rep(delta_right, 161)] 


data_train_DF<- data_train_DF %>% left_join(d %>% dplyr::select(starts_with("IdNumber")))
```

We can check

```{r}
sixselections <-sample(data_train_DF$IdNumber,6,rep=F)
ens=NULL
for(s in sixselections){ens=c(ens,data_train_DF %>% filter(IdNumber==s) %>% dplyr::select(starts_with("IdNum")) %>% as.vector() ) }
d<-data_frame(IdNumber=as.numeric(ens),coul=rep(1:6,each=5)) %>% left_join(data_train_DF %>% dplyr::select(IdNumber,lat,lon))
data_train_DF %>%  ggplot(aes(x=lon, y=lat))+
  geom_tile(color="black", fill="white")+geom_tile(data=d,fill=as.factor(d$coul))
```


## Time evolution

```{r}
data_train_DF %>% group_by(date,classe_CNT) %>% summarise(CNT=mean(CNT,na.rm=T)) %>%  ggplot(aes(x=date,y=log(CNT+1), color=as.factor(classe_CNT) ))+
  geom_line()+geom_smooth(se=0)+ theme(legend.position="none") +
  labs(title="Time evolution of Wildfires Counts", x="Date")
```


```{r}
data_train_DF %>% group_by(date,classe_BA) %>% summarise(BA=mean(BA,na.rm=T)) %>%  ggplot(aes(x=date,y=log(BA+1), color=as.factor(classe_BA) ))+
  geom_line()+geom_smooth(se=0)+ theme(legend.position="none") +
  labs(title="Time evolution of Wildfires Burnt areas", x="Date")
```
## Normal quantile transform

```{r}
order(c(1,NA,3,3,2))


```



